---
title: "Lab | Classes, functions & operators in R"
output: html_notebook
editor_options: 
  markdown: 
    wrap: 72
---

# Lab \| Classes, functions & operators in R

-   This lab is designed to help you practice creating custom classes,
    defining operators, and writing functions in R. Follow the
    instructions and complete the tasks below.

## Task 1: Create a Custom S3 Class

-   Define a custom S3 class called "Circle" that represents a circle.
    The class should store the radius of the circle.
-   Write a constructor function Circle() that takes the radius as input
    and assigns the class "Circle" to the object.

```{r}
Circle <- function(radius) {
  # Create a list to store the radius
  circle <- list(radius = radius)
  
  # Assign the class "Circle" to the object
  class(circle) <- "Circle"
  
  return(circle)
}
```

-   Write a print.Circle() method to display the radius and area of the
    circle when the object is printed. Use the formula for the area of a
    circle: area = π \* radius\^2.

```{r}
print.Circle <- function(circle) {
  # Calculate the area of the circle
  area <- pi * circle$radius^2

  # Print the formatted output
  cat("Type: Circle object\n")
  cat("Radius:", circle$radius, "\n")
  cat("Area:", area, "\n")
}
```

```{r}
# Create a Circle object with radius 5
my_circle <- Circle(5)

# Print the Circle object
print(my_circle)

```

Example: my_circle \<- Circle(5) print(my_circle) Output: Type: "Circle
object" <br> Radius: 5<br> Area: 78.53982<br>

## Task 2: Define a Custom Operator

1.  Define a custom operator %\>% that calculates the distance between
    the centers of two circles and checks if they intersect. The
    operator should return TRUE if the circles intersect and FALSE
    otherwise. The formula for the distance between two circles is:
    distance = sqrt((x2 - x1)\^2 + (y2 - y1)\^2).

    -   Assume that the circles are centered at (x1, y1) and (x2, y2),
        respectively.
    -   Two circles intersect if the distance between their centers is
        less than or equal to the sum of their radii.

```{r}
`%>%` <- function(circle1, circle2) {
  # Calculate the distance between the centers of the two circles
  distance <- sqrt((circle2$x - circle1$x)^2 + (circle2$y - circle1$y)^2)
  
  # Check if the circles intersect
  return(distance <= (circle1$radius + circle2$radius))
}
```

2.  Modify the Circle class to include x and y coordinates for the
    center of the circle. Example:

    ``` r
    circle1 <- Circle(radius = 3, x = 0, y = 0)
    circle2 <- Circle(radius = 4, x = 5, y = 0)
    circle1 %> circle2
    ```

    Output: TRUE

```{r}
Circle <- function(radius, x, y) {
  # Create a list to store the circle's properties
  circle <- list(radius = radius, x = x, y = y)
  
  # Assign the class "Circle" to the object
  class(circle) <- "Circle"
  
  return(circle)
}
```

3.  Example of Usage

```{r}
# Create two Circle objects with radius and center coordinates
circle1 <- Circle(radius = 3, x = 0, y = 0)
circle2 <- Circle(radius = 4, x = 5, y = 0)

# Use the custom operator to check if the circles intersect
intersecting <- circle1 %>% circle2

# Print the result
print(intersecting)
```

## Task 3: Write a Function for the Class

1.  Write a function circumference() that calculates the circumference
    of a Circle object. The formula for the circumference is:
    circumference = 2 \* π \* radius.

```{r}
# Define a generic circumference function
circumference <- function(object) {
  UseMethod("circumference")
}
```

```{r}
# Define a generic function
circumference <- function(object) {
  UseMethod("circumference")
}

# Define the method for the 'Circle' class
circumference.Circle <- function(object) {
  # Calculate the circumference using the formula
  circum <- 2 * pi * object$radius
  return(circum)
}
```

2.  Add this function as a method for the Circle class so that it can be
    called directly on objects of the class. Example: r my_circle \<-
    Circle(radius = 5) circumference(my_circle) Output: 31.41593

```{r}
# Define the method for the 'Circle' class
circumference.Circle <- function(object) {
  # Calculate the circumference
  circum <- 2 * pi * object$radius
  return(circum)
}
```

3.  Example of Usage

```{r}
# Assuming the Circle class is already defined to include x and y coordinates
my_circle <- Circle(radius = 5, x = 0, y = 0)

# Calculate the circumference of the circle
circumference_value <- circumference(my_circle)

# Print the result
print(circumference_value)
```

## Task 4: Create a Custom S4 Class

1.  Define a custom S4 class called "Rectangle" that represents a
    rectangle. The class should store the length and width of the
    rectangle.

```{r}
setClass(
  "Rectangle",
  slots = list(
    length = "numeric",
    width = "numeric"
  )
)
```

2.  Write a constructor function Rectangle() that takes the length and
    width as input and creates an object of the "Rectangle" class.

```{r}
Rectangle <- function(length, width) {
  # Construct a new "Rectangle" object
  new("Rectangle", length = length, width = width)
}
```

3.  Write a show() method for the Rectangle class to display the length,
    width, and area of the rectangle when the object is printed.

```{r}
setMethod(
  "show",
  "Rectangle",
  function(object) {
    # Calculate the area of the rectangle
    area <- object@length * object@width
    
    # Display the object's data
    cat("Type: Rectangle object\n")
    cat("Length:", object@length, "\n")
    cat("Width:", object@width, "\n")
    cat("Area:", area, "\n")
  }
)
```

```{r}
# Create a Rectangle object
my_rectangle <- Rectangle(length = 4, width = 6)

# Show the details of the Rectangle object
show(my_rectangle)
```

Example: r my_rectangle \<- Rectangle(length = 4, width = 6)
show(my_rectangle) Output: <br> Type: "Rectangle object" <br> Length: 4
<br> Width: 6 <br> Area: 24<br>

## Task 5: Define a Custom Operator for the S4 Class

1.  Define a custom operator %==% that compares two Rectangle objects
    and checks if they have the same area.

```{r}
# Define the custom operator for "Rectangle" class objects
setGeneric("%==%", function(rect1, rect2) standardGeneric("%==%"))

# Define the method for the operator
setMethod(
  "%==%",
  signature(rect1 = "Rectangle", rect2 = "Rectangle"),
  function(rect1, rect2) {
    # Calculate the areas of both rectangles
    area1 <- rect1@length * rect1@width
    area2 <- rect2@length * rect2@width
    
    # Return TRUE if areas are equal, otherwise FALSE
    return(area1 == area2)
  }
)
```

**Example Usage**

```{r}
# Create two Rectangle objects
rectangle1 <- Rectangle(length = 4, width = 6)
rectangle2 <- Rectangle(length = 3, width = 8)

# Use the custom operator to compare the areas
are_equal <- rectangle1 %==% rectangle2

# Print the result
print(are_equal)

```

2.  The operator should return TRUE if the areas are equal and FALSE
    otherwise.

Example: r rectangle1 \<- Rectangle(length = 4, width = 6) rectangle2
\<- Rectangle(length = 3, width = 8) rectangle1 %==% rectangle2 Output:
TRUE
